---
title: "Assigment2"
author: "Dannique Aalbu"
date: "May 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#Part 1: Importing & Tidying the California Gazetteer Dataset

## Read in a delimited text file ad saves it as a tibble

```{r Part1.1}
CAdata <- read_delim("CA_Features_20170401.txt", delim = "|")
gaz_raw <- as_tibble(CAdata)
```

## Formatting the tibble to include the desired columns and change the column names

```{r Part1.2}
gaz_tibble <- tibble(
  "Feature_ID" = gaz_raw$`<U+FEFF>FEATURE_ID`,
  "Feature_Name" = gaz_raw$FEATURE_NAME,
  "Feature_Class" = gaz_raw$FEATURE_CLASS,
  "State_Alpha" = gaz_raw$STATE_ALPHA,
  "County_Name" = gaz_raw$COUNTY_NAME,
  "Primary_Latitude" = gaz_raw$PRIM_LAT_DEC,
  "Primary_Longitude" = gaz_raw$PRIM_LONG_DEC,
  "Source_Latitude" = gaz_raw$SOURCE_LAT_DEC,
  "Source_Longitude" = gaz_raw$SOURCE_LONG_DEC,
  "Elevation_meters" = gaz_raw$ELEV_IN_M,
  "Map_Name" = gaz_raw$MAP_NAME,
  "Date_Created" = gaz_raw$DATE_CREATED,
  "Date_Edited" = gaz_raw$DATE_EDITED
)
```

##Changing the columns containing dates to a date format and convert any placeholders for unknown data to NA
```{r Part1.3}
gaz_tibble %>%
  mutate(Primary_Longitude = as.character(Primary_Longitude),Date_Created = as.Date(Date_Created, format = '%m/%d/%Y'),Date_Edited = as.Date(Date_Edited, format = '%m/%d/%Y')) 
gaz_tibble$Primary_Latitude <- parse_character(gaz_tibble$Primary_Latitude, na = "0")   
gaz_tibble$Map_Name <- parse_character(gaz_tibble$Map_Name, na = "Unknown") 
```

##Deleting rows where the primary latitude or longitude are unknown
```{r Part1.4}
gaz_tibble <- gaz_tibble %>%
  filter(Primary_Latitude != "NA") %>% 
  filter(State_Alpha == "CA")
View(gaz_tibble)
```

##Writing the gaz tibble to CSV file
```{r Part1.5}
write_delim(gaz_tibble, "GAZ.csv", delim = "|")
```

#Part 2: Data Analysis

##Question: what is the most commonly occurring feature name?
```{Part 2.1}
counter = count(gaz_tibble,Feature_Name,sort = TRUE)
common_feature <- which.max(counter$n)
counter[common_feature,c("Feature_Name")]
```

##Question: what is the least commonly occuring feature class?
```{Part 2.2}
counter2 = count(gaz_tibble,Feature_Class,sort = TRUE)
common_class <- which.min(counter2$n)
counter2[common_class,c("Feature_Class")]
```

##Question: what is the approximate center point of each county?
```{Part 2.3}
by_county <- gaz_tibble %>% 
  group_by(County_Name) %>%
  summarize(Min_lat = min(Primary_Latitude,na.rm = TRUE), Max_lat = max (Primary_Latitude, na.rm = TRUE), Min_long = min(Primary_Longitude, na.rm = TRUE), Max_long = max(Primary_Longitude, na.rm = TRUE)) %>%
  mutate(Min_lat = as.double(Min_lat), Min_long = as.double(Min_long), Max_lat = as.double(Max_lat), Max_long = as.double(Max_long)) %>%
mutate(xCenter = (Min_long + Max_long)/2) %>%
mutate(yCenter = (Min_lat + Max_lat)/2) 
#print the county names and the center point
print(subset(by_county, select = c(1, 6, 7)))
```

##Question: which features in each county are man-made?
```{Part 2.4}
by_features <- select(gaz_tibble,Feature_Class) %>%
mutate(Feature_Type = if_else(Feature_Class %in% c("Area","Arroyo","Bar","Bench","Bend","Park","Valley","Stream","Spring","Summit","Cape","Cave","Plain","Flat","Ridge","Woods","Beach","Lake","Bay","Channel","Island","Basin","Cliff","Arch","Pillar","Crater","Falls","Forest","Gap","Glacier","Gut","Isthmus","Lava","Range","Rapids","Sea","Slope","Swamp"),"Natural","Man-Made"))
print(by_features)
```